---
title: "MTH208: Worksheet 17"
format: 
  pdf:
    documentclass: article
editor: visual
geometry: margin=1in
header-includes:
   - \usepackage{amsmath, amssymb, setspace}
   - \onehalfspacing
   - \usepackage{etoolbox} 
   - \makeatletter 
   - \preto{\@verbatim}{\topsep=3pt \partopsep=3pt } 
   - \makeatother
---

# `C++` and `R` via `Rcpp`

By this time, you know that `R` (like many high-level languages) can be slower than expected. A lot of other high-level languages have this problem.

To make sure we can use R efficiently, we have a package called `Rcpp` that allows us to use `C++` from within R.

Let us see this by an example of adding two numbers:

```{r, cache = TRUE}
# Consider the following R function
addR <- function(x, y)
{
  return(x + y)
}

library(Rcpp)
# cppFunction() is a function that takes a bunch of C++ codes (all in quotes) 
# and then saves and compiles it.

# In C++, we have to declare the type of every object. 
# The function we are making is addC() which accepts two integers 
# and returns the sum of the two integers.
cppFunction('int addC(int x, int y) {
  int sum = x + y;
  return sum;
}')
```

When you run the above lines it takes some time for R to compile the function and get it ready. Once it has finished compiling, it makes available a function `addC` for you to use in `R`, that is actually written in `C++`.

```{r, cache = TRUE}
addR(3, 4)
addC(3, 4)
```

A typically `C++` code has the following form:

```{r}
#| eval: false

# for Rcpp usage put
# cppFunction('
output_type name_of_func(input_type input1, input_type input2, ...){

  line1 ... ;
  line2 ... ;
  .
  .
  return output;
# ')
}
```

Using Rcpp, we can also create functions that take vector inputs. Let's create functions that calculate the Euclidean distance between two vectors in both R and Rcpp.

```{r}
#| eval: false

# In R
EucR <- function(x, y)
{
  rtn <- sqrt(sum( (x-y)^2 )) 
  return(rtn)
}

# In C++ using Rcpp

cppFunction('double EucC(NumericVector x, NumericVector y) {
  double track = 0;
  int n = x.size();
  
  for(int i = 0; i < n; i++){
    track = track + pow( (x[i] - y[i]), 2);
  }
  track = sqrt(track);
  return track;
}
')
```

Please notice the change in syntax for loops etc. This is because inside `cppFunction` we are using `C++` compatible code. Run the following and make sure the output is TRUE:

```{r}
#| eval: false
x <- 1:10
y <- 3:12

# all.equal checks whether result is the same
all.equal(EucR(x, y),EucC(x, y))
```

To find out different kinds of data types in Rcpp, see this link here <https://teuder.github.io/rcpp4everyone_en/070_data_types.html>.

## Problems

1.  In the previous example, define `x` and `y` to be arbitrary vectors of length $10^4$. Benchmark the performance of `EucR` and `EucC`.\

2.  Consider a vector $x = (x_1, \dots, x_n)$, and suppose we want to calculate:

    $$
    \dfrac{\log(x_i)}{\sum_{k=1}^{n} \log(x_k)}
    $$

    The function `func` below calculates the above for a given vector `vec`.

    ```{r}
    #| eval: false

    func <- function(vec)
    {
      n <- length(vec)
      
      # for tracking sum and log
      sum.log <- 0
      log.of.vec <- numeric(length(n))
      
      # calculating logs and sum for each element
      for(i in 1:n)
      {
        log.of.vec[i] <- log(vec[i])
        sum.log <- sum.log + log.of.vec[i]
      }
      
      # fraction
      frac <- log.of.vec/sum.log
      return(frac)
    }
    ```

    i.  Rewrite the function in `R` in the most efficient way possible. Call this function `funcR`.

    ii. Rewrite the function using `Rcpp` and call the function `funcC`. Benchmark `func`, `funcR`, and `funcC`. Which is fastest?\

3.  Write an Rcpp function that adds two matrices together.\

4.  Writes an Rcpp equivalent of the function `colSums` and benchmark it against `colSums` for a large vector input.\

5.  Write an Rcpp function that takes a vector and returns a logical vector of the same length, indicating whether the value is positive (TRUE) or not (FALSE).\

6.  Redo Problem 4 in Worksheet 15 in Rcpp and benchmark with the other two implementations.\
