---
title: "MTH208: Worksheet 18"
format: 
  pdf:
    documentclass: article
editor: visual
geometry: margin=1in
header-includes:
   - \usepackage{amsmath, amssymb, setspace}
   - \onehalfspacing
   - \usepackage{etoolbox} 
   - \makeatletter 
   - \preto{\@verbatim}{\topsep=3pt \partopsep=3pt } 
   - \makeatother
---

## More `Rcpp`

Using the `cppFunction()` to integrate `C++` code in `R`, is very useful. However, it does not provide easy editing and indenting. Further, it is not practical to write big C++ codes in this way. A solution to this is `sourceCpp()` function.

Click on a New C++ file button, which creates a new file with C++ code that Rcpp can read. Save this file **in your working repository.**

The first two lines of the C++ file will be

```{r}
#| eval: false

#include <Rcpp.h>
using namespace Rcpp;
```

Note that the hashtag before include is necessary. Since this is a C++ file, \# is not a comment. Comments in C++ are using //.

Then, right before the function starts, you must write the following comment line to let `Rcpp` know that this is the main function:

```{r}
#| eval: false

// [[Rcpp::export]]
NumericVector dummy(NumericVector x){
  return x;
}
```

Once your file is written, in a separate `R` file, you can compile the `*.cpp` file using the `sourceCpp("*.cpp")` (this works because the file is in your working directory)

```{r}
#| eval: false
library(Rcpp)
sourceCpp("*.cpp") # replace * with name of file
```

Finally, you can learn about the various functions in Rcpp here: <https://teuder.github.io/rcpp4everyone_en/index.html>.

For generic C++ code see here: <https://www.w3schools.com/cpp/default.asp>

## Problems

1.  Redo Worksheet 17 problems using `sourceCpp()`.

2.  Using Section\~22 in the given Rcpp reference, redo Problem 3 in Worksheet 3. Benchmark with the solution using R.\
    The format for the while loop is:

    ```{r}
    #| eval: false
    int i;
    while(i < 10){
      ...
      i++;
    }
    ```

3.  Using Section\~22 in the given Rcpp reference, redo Problem 4 in Worksheet 3. Benchmark with the solution using R.\

4.  As you can see while dealing with matrices in C++, it can be annoying to write loops for elementary tasks like adding matrices. The `RcppArmadillo` library in R provides matrix compatibility.\
    \
    However, the namespace changes: `NumericVector` becomes `vec`, `NumericMatrix` becomes `mat`. For example, finding the inner product of two vectors, the cpp file looks like

    ```{r}
    #| eval: false
    // [[Rcpp::depends(RcppArmadillo)]]
    #include <RcppArmadillo.h>
    using namespace Rcpp;
    using namespace arma;

    // [[Rcpp::export]]
    double timesTwo(arma::vec x,arma::vec y) {
      // inner product of two vectors is a matrix
      mat inn(1,1); 
      inn = x.t() * y;
      
      // returning the first element of a 1x1 matrix
      return inn(0);
    }

    ```

    Using what you've learned here, write a C++ code using RcppArmadillo that adds two matrices and benchmark with the previous worksheet.\
    \
    More armadillo operations are here: <https://arma.sourceforge.net/docs.html>.
